### 2014-11-12 motivation.tex
edits to motivation

\noindent It takes a list of effects (the \emph{history} of updates to that
object), and an input argument, and returns a result along with an optional
effect (read-only operations return \cf{Nothing}). The new effect (if emitted)
is added to the state of the object at the current replica, and asynchronously
sent to other replicas. The implementation of the bank account operations in
\name is given in Figure~\ref{fig:ex}.

@@

The datatype \cf{Acc} represents the effect type for the bank account. The
function \cf{sum} returns the sum of elements in the list, and \cf{sel1}
returns the first element of a tuple. For each operation, \cf{hist} is a
\emph{snapshot} of the state of the object at some replica. In this sense,
every operation on the RDT is atomic, and thus permitting sequential reasoning
for implementing the RDTs. Here, \cf{getBalance} is a read-only operation,
\cf{deposit} always emits an effect, and \cf{withdraw} only emits an effect if
there is sufficient balance in the account. We have implemented a large corpus
of RDTs for realistic benchmarks including shopping carts, auction and
micro-blogging sites in few tens of lines of code.

@@

This notion of observable equivalence can be generalized to other RDTs as well.
For example, a last-writer-wins register with multiple updates is equivalent to
a register with only the last write. Similarly, a set with a collection of add
and remove operations is equivalent to a set with a series of additions of live
elements from the original set. Since the notion of observable equivalence is
specific to each RDT, we expect the programmer to provide a summarization
function (\cf{summarize}) of type \cf{[e] -> [e]} for each RDT, as a part of
the RDT specification.

The summarization function for the bank account is \cf{summarize hist =
[Deposit \$ sel1 \$ getBalance hist ()]}. Given a bank account history
\cf{hist}, the \cf{summarize} function returns a new history with a single
deposit of the current account balance. We utilize the summarization function
to ensure that the state of an object remains summarized in each replica.

>>>>

\noindent An operation takes a list of effects (the \emph{history} of updates to that
object), and an input argument, and returns a result along with an optional
effect (read-only operations return \cf{Nothing}). The new effect (if emitted)
is added to the state of the object at the current replica, and asynchronously
sent to other replicas. The implementation of the bank account operations in
\name is given in Figure~\ref{fig:ex}.

@@

The datatype \cf{Acc} represents the effect type for the bank account. The
function \cf{sum} returns the sum of elements in the list, and \cf{sel1}
returns the first element of a tuple. For each operation, \cf{hist} is a
\emph{snapshot} of the state of the object at some replica. In this sense,
every operation on the RDT is atomic, and thus amenable to sequential reasoning.
Here, \cf{getBalance} is a read-only operation,
\cf{deposit} always emits an effect, and \cf{withdraw} only emits an effect if
there is sufficient balance in the account. We have implemented a large corpus
of RDTs for realistic benchmarks including shopping carts, auction and
micro-blogging sites, etc. in a few tens of lines of code, expressed in this style.

@@

This notion of observable equivalence can be generalized to other RDTs as well.
For example, a last-writer-wins register with multiple updates is equivalent to
a register with only the last write. Similarly, a set with a collection of add
and remove operations is equivalent to a set with a series of additions of live
elements from the original set. Since the notion of observable equivalence is
specific to each RDT, programmers can provide a summarization
function - (\cf{summarize}) of type \cf{[e] -> [e]} -  as a part of
the RDT specification. The summarization function for the bank account is:

{\small
\begin{verbatim}
 summarize hist = 
   [Deposit $ sel1 $ getBalance hist ()]
\end{verbatim}}

\noindent Given a bank account history \cf{hist}, the \cf{summarize}
function returns a new history with a single deposit of the current
account balance. Our implementation invokes the summarization function
associated with an RDT to reduce the size of the effect sets
maintained by replicas.
